Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-03-15 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
upd

As we have seen in Sec.~\ref{sec:review_theories}, the QNM of slowly-rotating
BHs is modified gravity theories come in two families depending on their
transformation upon a parity transformation: axial and polar. Which is the one
we use to match against Eqs.~\eqref{eq:qnm_nongr_iso}?
%
To answer this question one has to work on a theory-by-theory case and in each
theory perform a translation between the metric perturbations $h_{\mu\nu}$ in
the Regger-Wheeler-Zerilli gauge and connect it with the transverse-traceless
gauge used to described GWs (see e.g.~Ref.~\cite{Maggiore:2018sht}, Sec. xxx),
%
In GR, both axial and polar QNMs are the same (a property known as \emph{isospectrality}) and therefore which
QNM we use to model the ringdown makes no difference.
%
In beyond-GR theories, isospectrality is in general broken (see the discussion
in Ref.~\cite{Hui:2021cpm} for a counterexample). and thus how axial and polar
gravitational QNMs appear in the GW signal has to be answered in
theory-by-theory.
%
This is outside the scope of this paper and here we take the more pragmatic
approach of simply choosing the \emph{least damped gravitational mode between
the two parities.}
%
The justification is this is the mode which, if excited during the merger, is
the most likely to appear in the signal.

We performed the mapping between theory-specific QNM calculation and the ParSpec framework
under the hypothesis above for the theories listed in Sec.~\ref{sec:review_theories}.
%
We leave the details of our calculation to Appendix~\ref{app:map_details} and
summarize our result in Table~\ref{tab:ref_theories_qnms}.

@@

In Fig.~\ref{fig:example_waveform} we show an illustrative waveform for the case of the cubic EFT
of gravity include the leading-order $n = 0$ deformations to the fundamental QNM.

@@

>>>>

As we have seen in Sec.~\ref{sec:review_theories}, the QNM of slowly-rotating
BHs is modified gravity theories come in two families depending on their
transformation upon a parity transformation: axial and polar. Which is the one
we use to match against Eqs.~\eqref{eq:qnm_nongr_iso}?
%
To answer this question one has to work with a chosen theory and perform a
translation between the metric perturbations $h_{\mu\nu}$ in the
Regger-Wheeler-Zerilli gauge and connect it with the transverse-traceless gauge
used to described GWs (see e.g.~Ref.~\cite{Maggiore:2018sht}, Sec. xxx),
%
In GR, both axial and polar QNMs are the same (a property known as \emph{isospectrality}) and therefore which
QNM we use to model the ringdown makes no difference.
%
In beyond-GR theories, isospectrality is in general broken (see the discussion
in Ref.~\cite{Hui:2021cpm} for a counterexample).
%
Thus, how axial and polar gravitational QNMs appear in the GW signal has to be
answered on theory-by-theory basis.
%
This is outside the scope of this paper and here we take the more pragmatic
approach of simply choosing the \emph{least damped gravitational mode between
the two parities.}
%
The justification is this is the mode (if excited during the merger) which is
the most likely to appear in the signal.

We performed the mapping between theory-specific QNM calculation and the ParSpec framework
under the hypothesis above for the theories listed in Sec.~\ref{sec:review_theories}.
%
We summarize our result in Table~\ref{tab:ref_theories_qnms} and
leave the details of our calculations to Appendix~\ref{app:map_details}.

@@

In Fig.~\ref{fig:example_waveform} we show an illustrative waveform for GR (solid line; using the \texttt{SEOBNR} model) and in the cubic EFT of gravity (dashed line; using the \pSEOB{} model),
including the leading-order $n = 0$ deformations to the fundamental QNM (see Table~\ref{tab:ref_theories_qnms}).
%
We chose binary parameters similar to GW150914, consisting of a quasicircular,
nonspinning binary with detector-frame masses
$m_1^{\rm dec} = 39\msun$ and $m_2^{\rm dec} = 31\msun$.
%
The top panel shows the $+$ polarization of the GW $h_{+}$ in both theories,
while the bottom panel shows the amplitude $|h| = \sqrt{h_{+}^2 + h_{\times}^2}$
and instantaneous frequency $f$, all as functions of time $t$.
%
The signal are identical up to the merger, after which they differ during the ringdown.
%
By construction, the ringdown is lasts longer for the cubic EFT of GR waveform and
with a smaller instantaneous frequency (see the bottom panel) due to the negative value of the
$\delta\omega^{(0)}$ coefficient in this theory.

@@

\hs{I suspect that by construction the $N_{\rm max} = 1$ will fail to satisfy our
perturbative bounds. If so, add explanation here.}
